Knowledge Representation for Service Robots: Supply and Demand

### 2012-10-08 krs-survey.tex
Merged from dissertation
-> many changes in many places

@@

The idea of \emph{Cognitive Robotics} was coined in the early 1990s by Reiter.
In a chapter on that subject in \emph{Fondations of Artifical
Intelligence}~\cite{Levesque2008}, Levesque reminds about the manifesto they
wrote together in 1998:

@@

        \item to execute a program, what information does a robot need to have
        at the outset vs. the information that it can acquire \emph{en route}
        by perceptual means?

        \item what does the robot need to know about its environment vs. what
        need only be known by the designer?

@@

\subsection*{A Note on Terminology}

Since this article discusses at length the concept of knowledge in the context
of robotics, it is useful to make our terminology explicit.  No general
agreement on a definition of ``knowledge'' exists. For our needs, we call
``knowledge'' \emph{an explicit set of interrelated logical facts that are
meaningful to the robot's executive controller}, where \emph{meaningful} means
that it can possibly be interpreted to lead to a purposeful action).

The relation of \emph{data} and \emph{information} to knowledge is a debated
epistemology question (the interested reader is invited to refer to the
Wikipedia page on the ``DIKW'' hierarchy). In this article, we will associate
\emph{data} to low-level material like raw sensor output, and
\emph{information} to uncontextualized symbolic facts.

To give a example, imagine a human looking at a book while being tracked by a
Kinect sensor: the pose of the human skeleton in the world would be the data,
the fact \concept{looksAt(entity1, entity2)} as computed by a geometric reasoning
module would be the information, the fact \concept{looksAt(john,
war\_and\_peace)}, fully grounded and connected to the whole knowledge base of
the robot would be proper knowledge.

@@

This exclude \textit{de-facto} platforms like {\sc DyKnow}~\cite{Heintz2004}
which are focused on data fusion and knowledge grounding at lower levels.

@@

While classical cognitive architectures like {\sc Soar}~\cite{Lehman2006} or
{\sc ACT-R} have declarative knowledge modules~\cite{Derbinsky2010} and have
been recently used on service robots (see {\sc ACT-R/E}~\cite{Kennedy2009} for
instance), they are also absent from this survey because we did not find much
references in the literature on knowledge manipulation and representation
applied to robotic scenarii for these architectures.

\subsection{Previous Work}
\label{sect|evaluation-literature}

A large part of our proposal is build as a comprehensive synthesis of
classification and analysis found in the literatture. This synthesis is focused
on cognitive abilities strictly related to knowledge manipulation, in the
context of service robotics.

Levesque and Lakemeyer~\cite{Levesque2008} present in a chapter on Cognitive
Robotics several characteristics of knowledge representation systems for
robots, stressing the need of representing the dynamics of the world.  Sensing
is included in the knowledge representation via fluents; they introduce the
idea of \emph{possible worlds} to represent distinct parallel mental models;
action representation and reasoning about tasks is discussed in the context of
situation calculus; open world vs. closed world approaches are both mentioned.
They also discuss how robot programming and knowledge representation can be
related. We integrate most of these items in our typology.

In a slightly broader context, Heintz et al.~\cite{Heintz2008} define
\emph{knowledge processing middlewares} as systems supporting ``declarative
specifications for flexible configuration and dynamic reconfiguration of
context dependent processing at many different levels of abstraction''. They
identify six characteristics: the system must be able to merge informations
from different, possibly distributed sources; it should support quantitative as
well as qualitative processing of information, it should offer bottom-up and
top-down processing, it should be able to deal with uncertainty, allow for
``flexible configuration and reconfiguration'' (which require what we call
\emph{non-monotonicity} in this article) and finally meta-knowledge and
introspective capacities (``declarative specification of the processing
functionalities''). In this paper, we focus our analysis on similar system with
however an emphasize on the symbolic capabilities.

Several surveys compare global cognitive architectures (\cite{Vernon2007, Chong2009}).
Vernon et al.~\cite{Vernon2007} split these architectures into two broad
categories: the \emph{cognitivist} ones (where cognition is considered as an
explicit computation problem, often based on symbol manipulation), and the
\emph{emergent} ones (where cognition only exists as a result of the
interaction of the system with its environment). The approaches presented in
this paper are, at a few exceptions, prototypical \emph{cognitivist} approaches
that aim at making knowledge explicit within the robot architecture. Vernon et
al. propose twelve \emph{characteristics of cognitive system} to compare
architectures. Amongst them, they mention the \emph{inter-agent epistemology} (how
the structure of the world is captured in a representation and shared), the
relation to \emph{embodiment}, the ability to \emph{anticipate} and to
\emph{adapt}, and the mechanisms of \emph{motivation}. While presented at the
level of the whole robotic architecture, these features also translate into
knowledge representation strategies and are relevant to our study.
\fxwarning{Detail Chong here}

@@

For instance, in~\cite{McCarthy2007}, McCarthy recently listed the challenges
he identifies on the road to a \emph{human-level AI}.

@@

	\item the ability to formalize and reason about contexts,

@@

Coming from the perspective of natural language processing in situated context,
Roy and Reiter, in~\cite{Roy2005}, summarize what they see as the main
challenges to be tackled by KRS: cross-modal representation systems,
association of words with perceptual and action categories (\emph{grounding}),
modeling of context, definition of the right granularity of models, integration
of temporal modeling and planning, ability to match past (learned) experiences
with the current interaction and ability to take into account the human
perspective.

Knowledge representation systems in robotics are concerned by most, if not all,
of these points, and we shall indeed mention them in our typology, in slightly
reformulated ways.

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.9\textwidth]{figs/brownie_scenario.jpg}
	\caption{A possible representation of the scenario, in the MORSE simulator}
	\label{fig|scenario}
\end{figure*}

In order to illustrate what we consider to be the required features of a
knowledge representation system for robotics, we introduce in this section an
(imaginary) demonstration scenario.

While devising a scenario inevitably constraints the breadth of abilities we
examine, we insist on the fact that our main purpose here is to ground into a
concrete case the set of representational features we consider as desirable.

We entitle our scenario ``the Brownie scenario'' (Figure~\ref{fig|scenario}):
Robi and Roba are two service robots, that can freely
move and pick objects around (with possibly different hardware and softwares
architectures, including different knowledge representation systems). They
cooperate with a human in a kitchen environment.

The main task of the scenario is the joint realization of a brownie, initiated
by the human: ``Let's make a brownie for tonight!''.

The scenario is successful if the task is achieved (the brownie is baked) in a
reasonable time (typically shorter than what it would have been required by the
human alone).

We voluntarily do not detail the subtasks of the scenario, neither we define
how they are shared amongst agents. In our analysis we focus on the general,
\textit{a priori} representational needs of the scenario.

A ``first-order'' analysis of this task leads to a rough partition of the
required representation abilities:

@@

% Representation of a complex task
We can further refine these categories: to prepare and bake a brownie, the
robot first needs to make sense of the term \emph{brownie} itself: what is it?
what is it used for? what is it made of? etc. We call this knowledge
\emph{common-sense knowledge} and the robot must be able not only to represent
it, but also to have access to an initial source (for instance through a initial
set of facts that are made available at startup, or via access to a Web-based
knowledge base like Wikipedia, etc.)

@@

These last assertions imply several other features for our knowledge
representation system: ``cooking often takes place in the kitchen'' implies
that representation of both uncertainty and likelihood is desirable. The fact
that cooking is associated to a place further implies that the system models
locations and is able to attach \emph{thematic relations} to concept
(here, the likely location of the cooking action).

``cooking requires ingredients'' hints about another very common feature
available in most knowledge representation systems: \emph{reasoning}. The robot
can \emph{infers} that cooking may require a recipe since a list of ingredients
is a pre-requisite of the cooking action, and a recipe may provide such a list.
If we omit the ``may'', this is a typical example of first-order logic
reasoning. Many other reasoning techniques exist (including probabilistic ones
-- ones able to deal with the ``may''), we shall illustrate some of them later
in this scenario.

We mentioned that a recipe often provides a procedure (or a \emph{plan}). The
robot should be able to store this plan in a way that allow later execution.
The plan is likely to contain \emph{spatio-temporal constraints} (like ``put
the brownie in the oven for 20 min'' or ``let's cook \emph{for tonight}'') that
must as well be appropriately represented.

To make decision, a robot may also want to \emph{predict} the state of the world
after some action (``if I leave the cake 2h in the oven, it will burn'').
Such ability to project itself in future or, generally speaking, in other
possible state of the world is related to several cognitive ability and
reasoning techniques: \emph{planning}, \emph{representation of possible worlds}
and \emph{non-monotonic reasoning}, in addition to common-sense knowledge and
\emph{physics-based} reasoning (that allows to predict that an egg is likely to
break if dropped).

Procedures are in addition often \emph{underspecified}: we can expect the recipe
to provide a cooking duration, but we usually do not expect the recipe to tell
us to first open the oven door, and then put the cake into it, since it is
self-evident that the door must first be opened to put the cake in the oven.
Such underspecification should be detectable, representable, and ideally
completable by the knowledge representation system\footnote{Note that we do not
assume here the {\it knowledge representation system} to be a single software
component: it may well be the result of the aggregation of several logical
components working together. We come back later on this aspect.}.

@@

Cooperation in our scenario can intervene at many places. For instance, an
agent may want to inform another one about the number of eggs that are
necessary for the brownie. This \emph{helping} behaviour makes sense only if
the first agent knows that the recipient agent both needs the information but
does not know it. This in turn requires the robot to be able to model the
knowledge of the other agents: to think \emph{from the perspective} of another
agent (an idea that is related to the availability of a theory of
mind~\cite{Leslie2000}).

Ability to communicate is one important pre-requisite to collaboration.
Communication in general requires the addresser and the addressee to share a
common interpretative framework (a shared common-sense knowledge -- or cultural
background -- and a shared context)~\cite{Jakobson1960}. In our scenario, the
agents are working in a kitchen. This element of context does not however
suffice if, for example, an agent asks another agent to ``give {[him]} the
bowl''. Behind the symbol ``bowl'', which physical entity are we actually
talking about? If we want to act on the world, this so-called \emph{grounding}
operation is essential, and may be tightly bound to the underlying knowledge
representation system. Note that grounding is a bidirectional process: in
covers the {\it top-down} operation (from the symbol to the percept) and the
{\it bottom-up} converse (retrieval or creation of symbols from perception).

A related ability is called \emph{pre-supposition accommodation}: if one of the
agent moves behind another one, with the brownie dough in its arm, and says
``be careful, I'm behind you!'', we want the first agent to be able to
represent both symbolically and geometrically (because, for instance, if the
agent want to move, it must take into account the new obstacle) something that
was not directly perceived. A knowledge representation system may be able to
provide structures to represent and reason about such cases.

Also central to cooperation are the notions of \emph{joint
intentions} and \emph{joint goals}~\cite{Tomasello2005, Bratman2009}: to help
the human during the cooking session, the robots need to track how far
they are into the recipe, what is the next step the human is likely to go for,
how task are currently split between agents, what action is currently
blocking the procedure, etc. This knowledge should let the robot identify the
intentions of other agents and create accordingly joint goals. Hence, a
knowledge representation system aiming at dealing with cooperative behaviours
is likely to have goal management structures taking explicitly into account
other agents' actions and goals.

In order to effectively share tasks, the robot must also know what it is
capable of: \emph{capability introspection} (both in term of general capability
and of immediate ability) is thus often desirable. More general introspection
(like the ability to tell ``who I am'' or ``what do I think of'') does not
appear to be necessary in our scenario. This may however be required in more
general human-robot interaction.

Last but not least, our scenario assumes implicitly \emph{natural interaction}
between humans and robots (as showed by the casual style of the
order ``Let's make a brownie!''). While natural language processing {\it
per-se} is usually out of the scope of a knowledge representation system, we
may want to be sure these systems may successfully interoperate, \ie that the
knowledge representation system provides efficient support to the natural
language processing module (for instance by adopting models and vocabulary that
are both well suited for machine processing and remain as close as possible to
the humans own structures and vocabulary).

@@

\subsubsection{Main Logic Formalisms}

The main role of a knowledge representation system is to provide an adequate
representation system to store facts and concepts that can be informally
described in natural language.

@@

Most (but not all) of the systems we survey rely on a particular logic
formalism. The choice of the formalism has a strong impact, on one side, on the
range of ideas that can be expressed conveniently (\emph{practical
expressiveness}) or at all (\emph{theoretical expressiveness}), on the other
side, on the ability to solve the inference problem (called
\emph{satisfiability}: is a given logical sentence true in my model?) in a
tractable manner.

A large number of logic formalism do exist, we shall summarize below the most
relevant ones for systems actually deployed in current robotic architectures.

@@

The family of \emph{Description Logics}~\cite{Baader2008} also play an
important role. It is also a subset of the first-order logic, with some
extensions in second-order logic. Description logics are notable because most
of them are known to be decidable (but not always in a practically tractable
manner). In description logic, axioms are build from \emph{concepts},
\emph{roles} (that are unary or binary predicates) and \emph{individuals}. The
W3C OWL-DL standard is a widely-used language to describe domains with the
description logic.

Because Description Logics have been originally created from the perspective of
a \emph{knowledge representation language} and not a logic language, their
terminology (\emph{concept} or \emph{class}, \emph{role} or \emph{property},
\emph{individual},\ldots) is well-suited to knowledge description and we may use
it in the remaining of this paper outside of the strict context of  Description
Logics.

\emph{Modal logic}, that allow for statement qualification like
\emph{possibility} or \emph{necessity}, have been shown to be closely related
to description logics~\cite{Baader2001}. Modal logic allows to represent conveniently parallel
possible worlds and facts like ``the robot knows \emph{that the human knows}
how to read a recipe''.

\fxfatal{On modal logics, see the remark of McCarthy, in \cite{McCarthy2007}, section 3}

\emph{Temporal logic} are designed to represent and manipulate assertions whose
truth value may vary in time.

@@

Note that most of these logic formalisms are still active research field on
their own, and practical considerations (especially the availability of
reasoners efficient enough for on-line use on a robot) often constraint the
choice of a logical formalism and a level of expressive power.

\subsubsection{Expressive Power}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.6\columnwidth]{expressive_overlap_dl_horn.pdf}
    \caption{expressiveness overlap of Description Logics and logic programs
    based on Horn clauses, taken from~\cite{Grosof2003}}
    \label{fig|overlap_dl_horn}
\end{figure}

Figure~\ref{fig|overlap_dl_horn} shows that the expressive power of description
logics and Horn clauses partially overlaps. In section~\ref{sect|reasoning} we
mention extensions to description logics based on rule systems to bring closer
the two approaches.

@@

\subsubsection{Open World and Close World Assumptions}
\label{sect|owa-cwa}

The \emph{close world} (CWA) vs. \emph{open world} (OWA) assumption names a
modelling choice on the \emph{completeness} of a knowledge domain. In the close
world assumption, a proposition that can not be proven true is assumed to be
false (\emph{negation by failure}), while in the open world assumption, a
proposition may be considered either true, false or unknown.

This distinction is important in robotics were the robot may have to manipulate
concepts with only partial knowledge on them. For instance, let imagine a robot
that sees a bottle on a table, whose bottom is hidden by another object. The
robot can not prove that the bottle is indeed \emph{on} the table. A knowledge
representation system relying on the closed world assumption would then assume
the bottle is \emph{not} on the table ($\lnot R^{CWA}_{isOn}(bottle, table)$)
whereas with the open world assumption, the proposition $R^{OWA}_{isOn}(bottle,
table)$ would be undecided. Example in table~\ref{table|cwa-owa-example} provides
another example of consequences of the CWA/OWA choice on reasoning.

@@

The OWL language is specifically known to assume an open world.  Domains
constrained with the closed world assumption lead to more tractable inference
problems, and allow for instance the use of logic languages like Prolog. Thus,
several approaches exists to \emph{locally close} a domain (\cf
Levesque~\cite{Levesque2008}, section 24.3.2 for a summary of those).

@@

@@

A knowledge representation system endowed with \emph{meta-cognition} is not
only able to manipulate knowledge but also to exhibit and manipulate the
structure of its knowledge and the reasoning process. For instance, the ability
to explain a logical inconsistency in a KRS is a meta-cognitive function.
Sloman~\cite{Sloman2011} \fxfatal{...Sloman?}

@@

\subsubsection{Role Representations}

This section discusses strategies and approaches to represent in a knowledge
model three important \emph{roles}\fxwarning{Reference on the term 'role'?}.

@@

Symbolic representation of space is a widely studied topic. In particular, a
large literature corpus is available on spatial ontologies.

@@

...topology??

@@

\paragraph{Representation of time}

>>>>

\fxnote{Support material: \emph{What is a knowledge representation} by Davis,
Shrobe and Szolovits,
\url{http://groups.csail.mit.edu/medg/ftp/psz/k-rep.html}}

@@

The idea of \emph{Cognitive Robotics} was coined in the early 1990s by Reiter.
In a chapter on that subject in \emph{Foundations of Artificial
Intelligence}~\cite{Levesque2008}, Levesque reminds about the manifesto they
wrote together in 1998:

@@

        \item to execute a program, what information does a robot need to have
        at the outset versus the information that it can acquire \emph{en route}
        by perceptual means?

        \item what does the robot need to know about its environment versus what
        need only be known by the designer?

@@

\paragraph{What do we call ``knowledge''?}
\label{sect|on-knowledge}

Since we will discuss at length the concept of knowledge in the context of
robotics in the coming pages, it is useful to make our terminology explicit.

Be it in philosophy, cognitive sciences or computer sciences, reaching an
agreement on a definition of ``knowledge'' seems difficult.

Allen Newell's famous \emph{Knowledge Level}~\cite{Newell1981} can be a
starting point: for Newell, \emph{knowledge} is a medium between \emph{agents}
and \emph{goals}, \emph{actions}, \emph{bodies}. Whereas the symbol level deals
with representation, the knowledge level deals with language, semantics;
whereas the symbol level deals with inference, the knowledge level deals with
entailment. We will, at the conclusion of the thesis, give a second look to
this distinction.

In our robotic context, we define knowledge as a narrower concept, while
keeping Newell's link to actions: ``knowledge'' is for us  \emph{a set of
interrelated logical facts that are meaningful to the robot executive
controller}. By \emph{meaningful} we mean that can possibly be interpreted to
lead to a purposeful action. We will see that our main challenge while
designing a cognitive architecture is furthermore to make this knowledge as
\emph{explicit} as possible.

The relation of \emph{data} and \emph{information} to knowledge is a debated
epistemology question known as the ``DIKW'' hierarchy question. In this thesis,
we will associate data to low-level material like raw sensor output, and
information to uncontextualised symbolic facts.

To give a example, we can imagine a human reading a book while being tracked by
a Kinect sensor: the pose of the human skeleton in the world would be the data,
the fact \concept{looksAt(human, book)} as interpreted by a geometric reasoning
module would be the information, the fact \concept{looksAt(john,
war\_and\_peace)}, fully grounded and connected to the whole knowledge base of
the robot would be proper knowledge.

@@

These criteria exclude platforms like {\sc DyKnow}~\cite{Heintz2004}
which are focused on data fusion and knowledge grounding at lower levels.

@@

While classical cognitive architectures like {\sc Soar}~\cite{Lehman2006},
GLAIR~\cite{Shapiro2009} or {\sc ACT-R} have declarative knowledge
modules~\cite{Derbinsky2010} and have been recently used on service robots (see
{\sc ACT-R/E}~\cite{Kennedy2009} for instance), they are also absent from this
survey because we did not find much references in the literature on knowledge
manipulation and representation applied to real-world robotic scenarii for
these architectures.

A comprehensive reference on (bio-inspired) cognitive architectures is also
available from the BICA Society~\cite{BicaCogArch2011}.

\subsection{Previous work}
\label{sect|evaluation-literature}

The typology we propose is in part based on a comprehensive synthesis of
classifications and analysis found in the literature. This synthesis is focused
on cognitive abilities strictly related to knowledge manipulation in the
context of service robotics.

Levesque and Lakemeyer~\cite{Levesque2008} present in their chapter on
Cognitive Robotics several characteristics of knowledge representation systems
for robots, stressing the need of representing the \emph{dynamics of the
world}.  Sensing is included in the knowledge representation via
\emph{fluents}; they introduce the idea of \emph{possible worlds} to represent
distinct parallel mental models; action representation and reasoning about
tasks is discussed in the context of \emph{situation calculus}; \emph{open
world} vs. \emph{closed world} approaches are mentioned.  They also discuss how
robot programming and knowledge representation can be related. We integrate
most of these items in our typology.

In a slightly broader context, Heintz et al.~\cite{Heintz2008} define
\emph{knowledge processing middleware} as systems supporting ``declarative
specifications for flexible configuration and dynamic reconfiguration of
context dependent processing at many different levels of abstraction''. They
identify six characteristics: the system must be able to \emph{merge
informations} from different, possibly distributed sources; it should support
quantitative as well as qualitative processing of information, it should offer
\emph{bottom-up} and \emph{top-down} processing, it should be able to deal with
\emph{uncertainty}, allow for ``flexible configuration and reconfiguration''
(which require what we call here \emph{non-monotonicity}) and finally
\emph{meta-knowledge} and \emph{introspective capacities} (``declarative
specification of the processing functionalities'').

Several surveys compare global cognitive architectures \cite{Langley2006,
Vernon2007, Chong2009}. Langley, Laird and Rogers~\cite{Langley2006}
distinguish nine capabilities: recognition and categorisation, decision making,
perception and situation assessment, prediction and monitoring, planning,
reasoning, execution control, interaction and
learning/remembering/introspection. They also separately identify four
\emph{properties} of a cognitive architecture, that categorise how knowledge is
handled by the architecture: representation of knowledge, organisation of
knowledge, utilisation of knowledge and acquisition and refinement of
knowledge. This categorisation had a notable influence on our typology, and
many of these categories are also present in our proposal.

Vernon et al.~\cite{Vernon2007} split these architectures into two broad
categories: the \emph{cognitivist} ones (where cognition is considered as an
explicit computation problem, often based on symbol manipulation), and the
\emph{emergent} ones (where cognition only exists as a result of the
interaction of the system with its environment). The approaches presented in
this paper are, at a few exceptions, prototypical \emph{cognitivist} approaches
that aim at making knowledge explicit within the robot architecture. Vernon et
al. propose twelve \emph{characteristics of cognitive system} to compare
architectures. Amongst them, they mention the \emph{inter-agent epistemology}
(how the structure of the world is captured in a representation and shared),
the relation to \emph{embodiment}, the ability to \emph{anticipate} and to
\emph{adapt}, and the mechanisms of \emph{motivation}. While presented at the
level of the whole robotic architecture, these features also translate into
knowledge representation strategies and are relevant to our study.

Chong et al.~\cite{Chong2009} also provide a recent review of the main cognitive
architectures, with a focus on eight functions: perception, memory, goals
management, problem solving capabilities, planning, reasoning, learning and
links to neurobiology.

@@

For instance McCarthy recently listed in~\cite{McCarthy2007} the challenges he
identifies on the road to a \emph{human-level AI}.

@@

	\item the ability to formalise and reason about contexts,

@@

Coming from the perspective of natural language processing in situated context,
Roy and Reiter summarise in~\cite{Roy2005} what they see as the main challenges
to be tackled by knowledge representation systems: \emph{cross-modal
representation} systems, association of words with perceptual and action
categories (\emph{grounding}), modeling of \emph{context}, definition of the
right \emph{granularity} of models, integration of \emph{temporal modeling and
planning}, ability to \emph{match past (learned) experiences} with the current
interaction and ability to take into account the \emph{human perspective}.

Knowledge representation systems in robotics are directly affected by these
points, and we indeed integrate them in our typology, in slightly reformulated
ways.

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.9\textwidth]{figs/brownie_scenario.jpg}
	\caption{The {\em Brownie} scenario}
	\label{fig|scenario}
\end{figure*}

The aim of this imaginary scenario (that has not been implemented, neither in
simulation nor on a robot) is to materialise early in this paper the context
and challenges of knowledge representation and manipulation for service, social
and interactive robotics. It underlines the place, the role and the need of
knowledge in a near-future, everyday situation where several robots and humans
co-exist and cooperate. This scenario will also be a source of support examples
for later sections of the thesis. Also, we have emphasised several keywords in
this description: we will come back to them in the next section to
explain them formally and relate them to each other.

We entitle our scenario ``the Brownie scenario'' (Figure~\ref{fig|scenario}):
Robi and Roba are two service robots, that can freely move and pick objects
around (with possibly different hardware and software architectures, including
different knowledge representation systems: typically, two robots built by two
different companies). They cooperate with a human in a kitchen environment.

The main task of the scenario is the joint realization of a brownie, initiated
by Tom, the human: ``Let's make a brownie for tonight!''.

The scenario is successful {\it 1)} if the task is achieved (the brownie is baked),
{\it 2)} in a reasonable time (typically shorter than what it would have been
required by the human alone), {\it 3)} and the human is happy by the help it received
(he/she did not feel useless, uncomfortable, unsafe).

We voluntarily do not detail the subtasks of the scenario, neither we define
how they are shared amongst agents: our focus is on knowledge needs and flows.

A ``first-order'' analysis of this task leads to a rough partition of the
required \emph{representation} abilities:

@@

% Representation of a complex task
We can further refine these categories: to prepare and bake a brownie, the
robot first needs to make sense of the term \emph{brownie} itself: what is it?
what is it used for? what is it made of? etc. We call this knowledge
\emph{common-sense knowledge} and the robot must be able not only to represent
it, but also to have access to an initial source (for instance through a
initial set of facts that are made available at startup, or via access to a
Web-based knowledge base like Wikipedia)

@@

These last assertions imply several other capabilities: ``cooking often takes
place in the kitchen'' implies that representation of both uncertainty and
likelihood is desirable. The fact that cooking is associated to a place further
implies that the system models locations and is able to attach \emph{thematic
relations} to concepts (here, the likely location of the cooking action).

``cooking requires a procedure and ingredients'' hints about another important
feature closely tied on knowledge manipulation: \emph{reasoning}. The robot can
\emph{infer} that cooking may require a recipe since a procedure and a list of
ingredients are pre-requisites of the cooking action that may be provided by a
recipe.  If we omit the ``may'', this is a typical example of first-order logic
reasoning.  Many other reasoning techniques exist (including probabilistic ones
-- ones able to deal with the ``may''), we shall illustrate some of them later
in this scenario.

We mentioned that a recipe often provides a procedure (or a \emph{plan}). The
robot should be able to store this plan in a way that allows later execution.
The plan is likely to contain \emph{spatio-temporal constraints} (like ``put
the brownie in the oven for 20 min'' or ``let's cook \emph{for tonight}'') that
must be as well appropriately handled.

To make decision, a robot may also want to \emph{predict} the state of the
world after some action (``if I leave the cake 2h in the oven, it will burn'').
Such ability to project itself in future or, generally speaking, in other
possible state of the world is related to several cognitive ability and
reasoning techniques: \emph{planning}, \emph{projection}, \emph{representation
of possible worlds} and \emph{non-monotonic reasoning}, in addition to
common-sense knowledge and \emph{physics-based} reasoning (that allows for
instance to predict that an egg is likely to break if dropped).

Procedures are in addition often \emph{underspecified}: we can expect the
recipe to provide a cooking duration, but we usually do not expect the recipe
to tell us to first open the oven door, and then put the cake into it, since it
is self-evident that the door must first be opened to put the cake in the oven.
Our cognitive robot should ideally be able to detect and possibly complete such
underspecification.

@@

Cooperation in our scenario can intervene at many places. For instance, an
agent may want to inform another one about the number of eggs that are
necessary for the brownie. This \emph{helping} behaviour makes sense only if
the first agent knows that the recipient agent both needs the information but
does not know it. This in turn requires the robot to be able to model the
knowledge of the other agents: to think \emph{from the perspective} of another
agent (an idea that is related to the availability of a theory of mind, we will
come back to it later on).

Ability to communicate is one important pre-requisite to collaboration.
Communication in general requires the addresser and the addressee to share a
common interpretative framework (a shared common-sense knowledge -- or cultural
background -- and a shared context). In our scenario, the agents are working in
a kitchen. This element of context does not however suffice if, for example,the 
human asks a robot to ``give {[him]} the bowl''. Behind the symbol
``bowl'', which physical entity are we actually talking about? If we want to
talk and act on the world, this so-called \emph{grounding} operation is
essential. It is a bidirectional process: in covers the {\it top-down}
operation (from the symbol to the percept) and the {\it bottom-up} converse
(retrieval or creation of symbols from perception).

A related ability is called \emph{pre-supposition accommodation}: if one of the
agent moves behind another one, with the brownie dough in its arm, and says
``be careful, I'm behind you!'', we want the first agent to be able to
represent both symbolically and geometrically (because, for instance, if the
agent want to move, it must take into account the new obstacle) something that
is not directly perceived.

Also central to cooperation are the notions of \emph{joint intentions} and
\emph{joint goals}: to help the human during the cooking session, the robots
need to track how far they are into the recipe, what is the next step the human
is likely to go for, how tasks are currently split between agents, what action
is currently blocking the procedure, etc. This knowledge should let the robot
identify the intentions of other agents and create accordingly joint goals.
Hence, a knowledge representation system aiming at dealing with cooperative
behaviours is likely to have goal management structures taking explicitly into
account other agents' actions and goals.

In order to effectively share tasks, the robot must also know what it is
capable of: \emph{capability introspection} (both in term of general capability
and of immediate ability) is thus often desirable. It can be extended to
general introspection (like the ability to tell ``who I am'' or ``what do I
think of'') that may be required for the interaction.

Last but not least, our scenario assumes implicitly \emph{natural interaction}
between humans and robots (as shown by the casual style of the order ``Let's
make a brownie!''), and we want to ensure that the
knowledge available to the robot provides efficient support to the natural
language understanding (for instance by adopting models and vocabulary that are
both well suited for machine processing and remain as close as possible to the
humans own structures and vocabulary), and also to \emph{non-verbal forms of
communication}, like gestures.

@@

This first axis of analysis is its intrinsic expressive power. It answers the
question: what can be possibly represented. When it explicitly exists, the
\emph{language} of representation plays here an obvious role.

\subsubsection{Main logic formalisms}

The main role of a knowledge representation system is to provide an adequate
representation system to formally store facts and concepts that could be informally
described in natural language.

@@

Most (but not all) of the systems we have surveyed rely on a particular logic
formalism. The choice of the formalism has a strong impact, on one side, on the
range of ideas that can be expressed conveniently (\emph{practical
expressiveness}) or at all (\emph{theoretical expressiveness}), on the other
side, on the ability to solve the fundamental inference problem (called
\emph{satisfiability}: is a given logical sentence true in my model?) in a
tractable manner.

A large number of logic formalisms do exist and we briefly present below the most
relevant ones for systems actually deployed in robotic architectures.

@@

The family of \emph{Description Logics}~\cite{Baader2008} also plays an
important role. It is also a subset of the first-order logic, with some
extensions in second-order logic. Description logics are notable because most
of them are known to be decidable (but not always in a practically tractable
manner). In description logic, axioms are build from \emph{concepts},
\emph{roles} (that are unary or binary predicates) and \emph{individuals}. The
W3C OWL-DL standard is a widely-used language to describe domains with the
description logic.

Because description logics have been originally created from the perspective of
a \emph{knowledge representation language} and not a logic language, their
terminology (\emph{concept} or \emph{class}, \emph{role} or \emph{property},
\emph{individual},\ldots) is well-suited to knowledge description.

\emph{Modal logic}, that allows for statement qualification like
\emph{possibility} or \emph{necessity}, have been shown to be closely related
to description logics~\cite{Baader2001}. Modal logic allows to represent conveniently parallel
possible worlds and facts like ``the robot knows \emph{that the human knows}
how to read a recipe''.

\fxwarning{On modal logics, see the remark of McCarthy, in \cite{McCarthy2007}, section 3}

\emph{Temporal logic} are designed to represent and manipulate assertions whose
truth value may vary in time. We introduce one of its key idea (the
\emph{fluents}) later in the typology.

@@

Note that most of these logic formalisms are still active research field on
their own, and practical considerations (especially the availability of
reasoners efficient enough for on-line use on a robot) often constrain the
choice of a logical formalism and a level of expressive power.

\subsubsection{Expressive power}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.3\columnwidth]{expressive_overlap_dl_horn.pdf}
    \caption{expressiveness overlap of Description Logics and logic programs
    based on Horn clauses, taken from~\cite{Grosof2003}}
    \label{fig|overlap_dl_horn}
\end{figure}

Figure~\ref{fig|overlap_dl_horn} shows that the expressive power of description
logics and Horn clauses partially overlaps. In section~\ref{sect|reasoning} we
mention extensions to description logics based on rule systems that bring closer
the two approaches.

@@

It can be noted that the relation between expressiveness and reasoning
complexity is fragile: for instance, adding the following axiom \stmt{farFrom
disjointProperty near} (that states that two individuals can not be at the same
time near and far from each other) changes the expressiveness power of the ORO
Common-Sense ontology (presented at chapter~\ref{chapt|oroserver}) from
$\mathcal{SHOIQ(D)}$ to $\mathcal{SROIQ(D)}$\footnote{See
appendix~\ref{chapt|dl} for a brief explanation of this notation}: this
seemingly innocuous assertion change the complexity class of the whole
ontology, and the concept satisfiability reasoning problem switches from a {\it
NExpTime-complete} problem to a {\it NExpTime-hard} problem (\ie, \emph{at
least} as hard as the hardest problem in {\it NExpTime}).

This ``instability'' has practical consequences on run-time performances on
the robot because a light alteration of the knowledge structure can lead to
very noticeable performance drops.

\subsubsection{Open world and close world assumptions}
\label{sect|owa-cwa}

The \emph{close world} (CWA) vs. \emph{open world} (OWA) assumptions name a
modelling choice on the \emph{completeness} of a knowledge domain. In the close
world assumption, a proposition that can not be proven true is assumed to be
false (\emph{negation by failure}), while in the open world assumption, a
proposition may be considered either true, false or unknown.

This distinction is important in robotics where the robot may have to manipulate
concepts with only partial knowledge on them. For instance, let us imagine a robot
that sees a bottle on a table, whose bottom is hidden by another object. The
robot can not prove that the bottle is indeed \emph{on} the table. A knowledge
representation system relying on the closed world assumption would then assume
the bottle is \emph{not} on the table ($\lnot R^{CWA}_{isOn}(bottle, table)$)
whereas with the open world assumption, the proposition $R^{OWA}_{isOn}(bottle,
table)$ would be undecided. Example in table~\ref{table|cwa-owa-example} provides
another example of consequences of the CWA/OWA choice on reasoning.

@@

The OWL language is specifically known to assume an open world.  Domains
constrained with the closed world assumption lead to more tractable inference
problems, and allow for instance the use of logic languages like Prolog. Thus,
several approaches exists to \emph{locally close} a domain (\cf
Levesque~\cite{Levesque2008}, section 24.3.2 for a summary of those ones).

@@

\fxfatal{Representation of uncertainty: To be completed!}

@@

Sloman proposes in~\cite{Sloman2011} a detailed analysis of meta-cognition and
its different aspects in both natural (human) and artificial systems.

A knowledge representation system endowed with \emph{meta-cognition} is not
only able to manipulate knowledge but also to exhibit and manipulate the
structure of its knowledge and the reasoning process. For instance, the ability
to explain a logical inconsistency in a KRS is a meta-cognitive function, as is
the ability to expose and alter the knowledge structure (these two reasoning
techniques have their own entries in the taxonomy, at
section~\ref{sect|reasoning}.

@@

\subsubsection{Role representations}

This section discusses strategies and approaches to represent in a knowledge
model three important \emph{roles}: spatial relations, time and representation
of actions.

@@

Symbolic representation of space is a widely studied topic. In particular, a
large literature corpus is available on spatial ontologies\fxwarning{Which one?}.

@@

Topological maps are abstractions of an environment as graphs where nodes
represent places and edges represent connections between places. Because they
are symbolic representation, topological maps allow higher-level reasoning
(such as containment, connectivity, regions) than metric maps.

One important contribution to the building of a coherent representational stack
for space representation is the Spatial Semantic Hierarchy~\cite{Kuipers2000}
introduced by Kuipers. It consists in multiple interacting representations of
space, both qualitative and quantitative, that span from sensor-level
representation to ontologies of places and regions.

@@

\paragraph{Representation of time}
\label{sect|time-representation}
